Future challenges for Research Software Engineering

### 2025-05-09 future_challenges_for_RSEng.tex
Impact on structures

\subsection{Impact of Alternate Career Paths for Researchers on the Structure of Research Institutes}
The RSE is probably one of the first professionalizations in the academic mid-level that is able
to collectively voice their concerns. This enables the possibility to reshape universities.
If more people are employed as mid-level, this makes a steadier pool of people
available that can integrate new people that are regularly coming into departments and know the local facilities.
On the other hand, an upcoming challenge how universities decide to design the contracts of people employed
in RSE units. Will they be permanent, or will there also be an up-or-out mentality?

As outlined in [cite RSE group paper] not all universities need to opt for strengthening
the academic mid-level. Some might have chosen to set up external companies to outsource typical RSE tasks,
thereby providing career paths outside of academia.

FG-NOTE: By Nick. I've put it here, due to the relation with the RTP question.

>>>>

\subsection{Impact of Alternate Career Paths for Researchers on the Structure of Research Institutes}

RSEs can either be embedded in research groups (belonging to the mid-level university staff),
or be part of larger RSE units.
The participants had the opinion that focusing on embedded RSEs would make onboarding new RSEs to existing projects easier.
This, however, comes with additional funding challenges,
as long-term funding is more difficult to acquire in lower levels of an organization.
Third-party funding can mitigate this issue in the mid-term.
However, the participants considered central RSE groups more likely in the future,
due to the fact that there are already examples and activities from various universities in this direction
(examples include the univerisities of Jena, Heidelberg, Göttingen).
We expect funding in central groups to evolve from initially temporary positions to permanent positions,
as the groups themselves prove themselves valuable.
Embedded RSEs and central groups will need to network with each other,
and potentially request and coordinate support from external RSEs.
